---
title: "Data Wrangling - Part II"
subtitle: "StatPREP R Workshops"
author: "Onalee Garusinghe"
output: html_document
---

```{r setup, include=FALSE}
# Don't forget to load your libraries
library(tidyverse)
library(ggplot2movies)
library(ggformula)
```

```{r}
# This loads the data from the ggplot2movies package
data(movies)
```

# Deriving information

So far, you have learned "verbs" for isolating data, `select`, `filter`, `arrange`. Now, we will learn verbs for deriving information, `summarize`, and `mutate`. There is an additional verb, `group_by`, which is useful in conjunction with these verbs, but not as much on its own. 

## summarize

Summarize is a many-to-one or many-to-few operation. It summarizes the dataset down to fewer numbers. 

Let's say we want to know the average budget of the movies in the `movies` dataset. The way to compute that is with summarize,

```{r}
movies %>%
  summarize(mean(budget))
```

Unfortunately, there are many missing (or Not Available) values in `budget`, so we need to remove those in our pipeline.

```{r}
movies %>%
  drop_na(budget) %>%
  summarize(mean(budget))
```

Notice that this pipeline does not change the `movies` dataset in our environment. The NA values are still present in the data, and the entire dataset is just how it started.  We just `summarize`d the data down to a single number in our Console or RMarkdown document. 

If we want, we can add additional statistics to our `summarize` command,

```{r}
movies %>%
  drop_na(budget) %>%
  summarize(mean(budget), max(budget))
```

We can also name the statistics we compute, so we could refer to them later.

```{r}
movies %>%
  drop_na(budget) %>%
  summarize(avg_budget = mean(budget), max_budget = max(budget))
```
#Question 1
### Your turn

Now, you try it! 

- What year was the earliest movie in the dataset released? 
- What is the average length of movies?
- What is the standard deviation of length?

You may want to look at the "data transformation with dplyr" cheatsheet to review some possible summary functions. 

```{r eval=TRUE}
movies_lin71<-movies %>%
  summarize(mean(length),sd(length),min(year))
movies_lin71
```

The oldest movies on this list are movies made in 1893. The average length of movies is 82 minutes and about 34 seconds. Although there is a standard deviation of 44 minutes and about 35 seconds. 

#Question 2
## group_by

The `group_by` function groups cases by common values. 

This is useful if want to compute summary statistics by group. For example, we may be interested in whether action movies are longer than non-action movies. 

We already know how to compute the average length of movies overall. What we need to add to our pipeline is a `group_by`

```{r eval=TRUE}
movies_ln84<-movies %>%
  group_by(Action) %>%
  summarize(mean(length),sd(length),min(year))
  # move your summarize code from the last exercise here!
movies_ln84
```

Which movies are longer, action movies or non-action movies? 

On average Action movies are seen to be longer with run times of about 100 minutes compared to non-action movies having run times of about 81 minutes. The standard deviation also shows the difference between the two as it is lower with the Action movies eaning more the movies computed are  closer to 100 minutes while non-action movies vary more on run time.

#Question 3
### Your turn

Investigate the movies classified as Short. Are they actually shorter than other movies? 

- What is the average length? 
- What is the minimum length?

```{r eval=TRUE}
movies_ln99<-movies %>%
  group_by(Short) %>%
  summarize(mean(length),min(length))
movies_ln99
```

Short movies have a average length of close to 14 minutes, while non-short movies have an average length of about 95 minutes. Interestingly enough, Short and non-short movies have the same lowest run time for lowest length of a movie, both being 1 minute as the minimum length of a movie.

## mutate

The `mutate` function creates new variables (columns) in the dataset. Typically, we use it to make new variables from old ones. 

For example, we may be interested in the length of movies in hours, rather than in minutes. 

```{r}
movies <- movies %>%
  mutate(length_hrs = length / 60)
```

#Question 4

Let's read this code carefully. 

- what does the `<-` do?
- why are we writing `/60`?
- what does the `=` do?
- why did we use the name `length_hrs`?

  The "<-" operator is used to assigns the value on the right to the variable on the left. The "/60" is typically used to convert minutes into hours, as there are 60 minutes in an hour. The Role of "=" is also used to assign the value of the correct answer or value on the right to the variable on the left, and last the name "length_hrs" likely indicates that the variable represents a length of time measured in hours, providing clarity and context within the code.

###

The variables `r1`, `r2`, ..., `r10` give you the percent of people who rated that movie a 1, 2, ..., 10. We would like to know the percent of people who rated the movie a 4 or less (i.e., who rated it a 1, 2, 3, or 4). Create a new variable called `per_low` that computes that number.

```{r eval=TRUE}
movies_ln128 <- movies %>%
  mutate(per_low = r1+r2+r3+r4)
movies_ln128
```

#Question 5

Now that we have a bunch of wrangling skills, let's put them together. 
- pull out the movies that had a high percent of low ratings (Hint: what command will you use to do this?)
- pull out movies that had a high percent of low ratings *and* more than 5 votes. 
- how many of these bad movies were there?

To accomplish that last task, we need a specialty function, `n`. Here's how it works

```{r}
movies_ln144<-movies_ln128 %>%
  filter(per_low > 50 & votes > 5)%>%
  select(per_low,votes,title) %>%
  summarize(Bdmvs = n())
movies_ln144
```

4,753 movies have a high percent of low ratings with a vote greater than 5.

- what is the average length of movies in the dataset, in hours? 
- can you find a second way to compute the average length, by flipping the order of commands (i.e., `summarize` first and then `mutate`)?

